Binary Latent Representations for Efficient Ranking

### 2017-06-07 paper/paper.tex
Notation improvements, latex table output.

@@

\documentclass[sigconf]{acmart}
\begin{document}

\begin{abstract}
  "Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum."
\end{abstract}

@@

To alleviate these constraints, we propose the use of binary item and user representations in learning-to-rank matrix factorization models. Following the approach of \cite{rastegari2016xnor}, we show that we can estimate binary user and item representations that achieve comparable accuracy to standard learning-to-rank approaches while being several times faster to score and requiring a fraction of the memory. In our MovieLens 100k experiments, our binary representations are within 1\% of the accuracy of the real-valued model, but are over 2 times faster to score and take 50\% of the memory.

\section{Related work}
\subsection{Heuristics}
\subsection{Approximate inner product search}
\cite{shrivastava2014asymmetric} propose a solution fo efficient ranking of large result sets through approximate maximum inner product search (MIPS). They extend well-known \citep{indyk1998approximate} locality-sensitive hashing techniques to the MIPS setting through asymmtric transformations of the query and candidate vectors, and show their applicability to recommender systems.

@@

\cite{Shen_2015_ICCV} expand on this approach by learning asymmetric binary hash functions. The hash functions are trained to minimize the L2 norm between the inner product matrix of the original input vectors and their trained binary representations.
\subsection{Quantization}

\section{Our approach}
\cite{rastegari2016xnor} introduce XNOR-Networks, an approach that combines 1-bit quantization of fully-connected and convolutional layers with representing dot products through scaled XNOR and bitcounting operations.

@@

In a traditional real-valued model, the prediction $r_{ui}$ for any user-item interaction pair $(u, i) \in \symUserSet \times \symItemSet$ would be given by
\begin{equation}
r_{ui} = \sigma(\vec{u_i}) \cdot \vec{i_i} + b_i + b_u)
\end{equation}
where $\sigma$ denotes the sigmoid function, $\vec{u_i}$ the user and $\vec{i_i}$ the item latent vector.

In the binary representation model the above is approximated by
\begin{equation}
\mathrm{sign}(\vec{u_i}) \cdot \mathrm{sign}(\vec{i_i}) \cdot \Vert\vec{u_i}\Vert_{l1} \cdot \Vert\vec{i_i}\Vert_{l1}
\end{equation}

@@

We train the model using the Bayesian Personalised Ranking (BPR, \cite{rendle2009bpr}) loss,

@@

We use the approach developed by XNOR-Net to approximate latent-space dot products through binary operations. In this approach, we approximate the dot product between the user vector $\vec{u_i}$ and the item vector $\vec{i_i}$ by
\begin{equation}
\mathrm{sign}(\vec{u_i}) \cdot \mathrm{sign}(\vec{i_i}) \cdot \Vert\vec{u_i}\Vert_{l1} \cdot \Vert\vec{i_i}\Vert_{l1}
\end{equation}

@@

Benchmark results are measured on an Intel Xeon E5-2686v4 CPU.

@@

In the binary version, the 1-bit weights are packed into 32-bit integer buffers. The XNOR dot product is implemented using 8-integer wide XOR operations (256 binary weights are processed at a time), followed by a popcount instruction to count the number of on bits in the result. We use the \texttt{libpopcnt}\footnote{\texttt{https://github.com/kimwalisch/libpopcnt}} \citep{mula2016faster} library for the bit counting operations.

@@

The code is compiled using GCC 5.4.0 for the AVX2-enabled Broadwell architecture.

\subsection{Results}

\bibliographystyle{abbrv}
\bibliography{bibliography}

>>>>

\documentclass[sigconf]{acmart}

\usepackage{booktabs}
\usepackage{commath}

%\usepackage{tabularx}

@@

\DeclareMathOperator{\sign}{sign}
\renewcommand\vec{\mathbf}

\begin{document}

\begin{abstract}
Multifrequency media access control has been well understood in
general wireless ad hoc networks, while in wireless sensor networks,
researchers still focus on single frequency solutions. In wireless
sensor networks, each device is typically equipped with a single
radio transceiver and applications adopt much smaller packet sizes
compared to those in general wireless ad hoc networks. Hence, the
multifrequency MAC protocols proposed for general wireless ad hoc
networks are not suitable for wireless sensor network applications,
which we further demonstrate through our simulation experiments. In
this article, we propose MMSN, which takes advantage of
multifrequency availability while, at the same time, takes into
consideration the restrictions of wireless sensor networks. Through
extensive experiments, MMSN exhibits the prominent ability to utilize
parallel transmissions among neighboring nodes. When multiple physical
frequencies are available, it also achieves increased energy
efficiency, demonstrating the ability to work against radio
interference and the tolerance to a wide range of measured time
synchronization errors.
\end{abstract}

@@

To alleviate these constraints, we propose the use of binary item and user representations in learning-to-rank matrix factorization models. Following the approach of \citet{rastegari2016xnor}, we show that we can estimate binary user and item representations that achieve comparable accuracy to standard learning-to-rank approaches while being several times faster to score and requiring a fraction of the memory. In our MovieLens 100k experiments, our binary representations are within 1\% of the accuracy of the real-valued model, but are over 2 times faster to score and take 50\% of the memory.

\section{Related work}
\citet{shrivastava2014asymmetric} propose a solution fo efficient ranking of large result sets through approximate maximum inner product search (MIPS). They extend well-known \citep{indyk1998approximate} locality-sensitive hashing techniques to the MIPS setting through asymmtric transformations of the query and candidate vectors, and show their applicability to recommender systems.

@@

\citet{Shen_2015_ICCV} expand on this approach by learning asymmetric binary hash functions. The hash functions are trained to minimize the L2 norm between the inner product matrix of the original input vectors and the inner product matrix of their binary representations. As a post-processing technique, it is a applicable to a wide range of problems. However, because the binary representations are obtained as a separate step, this approach may not be able to exploit the benefits in training a binary representation approach end-to-end.

\section{Our approach}
\citet{rastegari2016xnor} introduce XNOR-Networks, an approach that combines 1-bit quantization of fully-connected and convolutional layers with representing dot products through scaled XNOR and bitcounting operations.

@@

In a traditional real-valued latent factor model, the prediction $r_{ui}$ for any user-item interaction pair $(u, i) \in \symUserSet \times \symItemSet$ is given by
\begin{equation}
r_{ui} = \sigma(\vec{u}_i \cdot \vec{i}_i + b_i + b_u)
\end{equation}
where $\sigma$ denotes the sigmoid function, $\vec{u}_i$ the user and $\vec{i}_i$ the item latent vectors, and $b_u$ and $b_i$ the user and item biases.

\citet{rastegari2016xnor} show that the dot product of two real-valued vectors can be approximated in the binary domain by (1) binarizing the input vectors using the sign function (so that the results lie in $\{1, -1\}$), (2) taking their dot product, and (3) scaling the result by the average magnitude of the vectors' elements:
\begin{equation}
\vec{u}_i \cdot \vec{i}_i \approx \sign(\vec{u}_i) \cdot \sign(\vec{i}_i) \cdot \sum_k\abs{u_{ik}} \cdot \sum_k\abs{i_{ik}}
\end{equation}

Applying this to the latent-factor model, the prediction for a user-item pair is given by
\begin{equation}
r_{ui} = \sigma\left(\sign(\vec{u}_i) \cdot \sign(\vec{i}_i) \cdot \sum_k\abs{u_{ik}} \cdot \sum_k\abs{i_{ik}} + b_i + b_u\right)
\end{equation}

@@

We train the model using the Bayesian Personalised Ranking (BPR, \citet{rendle2009bpr}) loss,

@@

@@

Benchmark results are measured on an Intel Xeon E5-2686v4 CPU by running a 100 repetitions of scoring 100,000 items and averaging the results.

@@

In the binary version, the 1-bit weights are packed into 32-bit integer buffers. The XNOR dot product is implemented using 8-integer wide XOR operations (256 binary weights are processed at a time), followed by a popcount instruction to count the number of on bits in the result. We use the \texttt{libpopcnt} \citep{mula2016faster} library for the bit counting operations.

@@

The code is compiled using GCC 4.8.4 for the AVX2-enabled Broadwell architecture.

\subsection{Results}
\begin{table*}
\centering
\begin{tabular}{rrrrrrrr}
\toprule
Dimension &   MRR & Binary MRR & MRR ratio &   PPMS & Binary PPMS & PPMS ratio & Memory use ratio \\
\midrule
       32 & 0.096 &      0.088 &     0.919 & 89,299 &     245,644 &      2.751 &            0.091 \\
       64 & 0.098 &      0.096 &     0.973 & 47,565 &     209,486 &      4.404 &            0.062 \\
      128 & 0.108 &      0.091 &     0.843 & 20,502 &     156,404 &      7.629 &            0.047 \\
      256 & 0.109 &      0.097 &     0.884 & 10,631 &     166,500 &     15.661 &            0.039 \\
      512 & 0.112 &      0.105 &     0.938 &   5391 &     110,418 &     20.479 &            0.035 \\
     1024 & 0.114 &      0.102 &     0.897 &   2689 &      69,349 &     25.786 &            0.033 \\
     2048 & 0.111 &      0.109 &     0.984 &   1352 &      34,826 &     25.745 &            0.032 \\
     4096 & 0.111 &      0.107 &     0.961 &    666 &      15,907 &     23.856 &            0.032 \\
\bottomrule
\end{tabular}
\end{table*}

\bibliography{bibliography}
\bibliographystyle{ACM-Reference-Format}
